Language-Based Security for Low-Level MPC

### 2024-05-14 ppdp24/bruteforce.tex
checkpoint

In $\mathbb{F}_2$ a brute force strategy for computing
$\progtt(\prog)$ for any $\prog$ is to directly compute
$\runs(\prog)$. By querying $\progtt(\prog)$ we can verify any of the
hyperproperties discussed previously, or other properties such as
perfect secrecy \cite{XXX}.  Since $\runs(\prog)$ is exponential in
the size of $\secrets(\prog) \cup \flips(\prog)$, this strategy is
feasible only for smaller protocols. However, we can efficiently
convert protocols in the passive setting (without $\ttt{assert}$) to
stratified Datalog programs, and then extract $\runs(\prog)$ by
calculating Least Herbrand models by parallelization and/or other HPC
acceleration techniques for logic programs \cite{aspis2018linear}.
The rewriting we describe here is to Datalog with negation, with a
negation-as-failure model, though we can also use techniques in
\cite{sakama2017linear} to eliminate negation from resulting programs

@@

\begin{comment}
We immediately note that $\runs(\prog)$ can be obtained by a left-folding
of $\solvealg$ across $\prog$. 
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\xassign{x}{\be}{\cid}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
  \end{eqnarray*}
  Then $\cruns(\prog) = \runs(\prog)$.
\end{lemma}
\end{comment}

@@

  \vars\ (\be_1 \exor \be_2)\ \cid \defeq (\vars\ \be_1\ \cid) \cup (\vars\ \be_2\ \cid)
\end{mathpar}
... and so on. Then, to begin converting a protocol $\prot$ to a Datalog
we define the function ${tt}$ that applies $\solve{\!}{\!}{\!\!}$ locally to each command
in $\prog$, obtaining the subset of memories that result
in a variable assignment of $1$.  
\begin{definition} Define:
$$
{tt}(\xassign{x}{\be}{\cid}) \defeq (x, (\solve{(\mems(\vars\ \be\ \cid))}{\be}{\cid}))
$$
\end{definition}
The mapping of ${tt}$ across a program
$\prog$-- i.e., $(\mathit{map}\ {tt}\ \prog)$--  essentially defines the
logic program, modulo some syntactic conversion. We can
accomplish the latter as follows, where $\datalog(\prog)$ defines the
full conversion.
\begin{definition} We define the conversion from memories to
  literals and clause bodies as follows:
\begin{mathpar}
  \logit{x \mapsto 1} \defeq x

@@

  \logit{\{ x_1 \mapsto \beta_1, \ldots, x_n \mapsto \beta_n\}} \defeq
  \logit{x_1 \mapsto \beta_1} \wedge \cdots \wedge \logit{x_n \mapsto \beta_n}
\end{mathpar}
Given pairs $(x,\stores)$ in the range of ${tt}$, we define the conversion
to clauses as  $\mathit{clauses}(x,\{ \store_1,...,\store_n \}) \defeq x \gets \logit{\store_1} \vee \cdots \vee x \gets \logit{\store_n}$.
The $\minifed$-to-Datalog conversion is then defined as:
$$
\datalog(\prog) \defeq  \mathit{map}\ \mathit{clauses}\ (\mathit{map}\ {tt}\ \prog)
$$
\end{definition}

@@

\paragraph{Verifying Security Properties} We can
query correct representations of $\runs(\prog)$ by using
implementations of conditioning and marginalization to automatically
verify the passive model hyperproperties described in Section
\ref{section-hyper}-- in particular correctness and noninterference
modulo output, which imply passive security.  We have used this method
to verify security in $\mathbb{F}_2$ of Shamir addition as defined in Section
\ref{section-lang-example}, and single-gate $\eand$ and $\exor$ circuits with GMW
Beaver triples. We have also
verified properties compositional properties of gates themselves that establish
circuit invariants that imply passive security in larger circuits.
We discuss these examples and proof methods in Section
\ref{section-examples}.

>>>>

In $\mathbb{F}_2$ a brute force strategy for computing
$\progtt(\prog)$ for any $\prog$ is to directly compute
$\runs(\prog)$. By querying $\progtt(\prog)$ we can verify any of the
hyperproperties discussed previously, or other properties such as
perfect secrecy \cite{XXX}.  A basic method to do this is
to calculate a truth table for the given protocol. Since
$\runs(\prog)$ is exponential in
the size of $\secrets(\prog) \cup \flips(\prog)$, this strategy is
feasible only for smaller protocols. However, to support some
scaling we can efficiently
convert protocols in the passive setting (without $\ttt{assert}$) to
stratified Datalog programs, and then extract $\runs(\prog)$ by
calculating Least Herbrand models by parallelization and/or other HPC
acceleration techniques for logic programs \cite{aspis2018linear}.
The rewriting we describe here is to Datalog with negation, with a
negation-as-failure model, though we can also use techniques in
\cite{sakama2017linear} to eliminate negation from resulting programs.

\subsection{Computing a Truth Table}

@@

We immediately note that $\runs(\prog)$ can be obtained by a left-folding
of $\solvealg$ across $\prog$. 
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\xassign{x}{\be}{\cid}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}%\\[2mm]
    %\cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup R)\ \prog
  \end{eqnarray*}
  Then assuming default preprocessing, $\mathit{foldl}\ {tt}\ \mems(S \cup R)\ \prog = \runs(\prog)$.
\end{lemma}
However, this method does not take advantage of parallellization,
in that elements of $\runs(\prog)$ can be calculated independently.

\subsection{Conversion to Stratified Datalog}

@@

  \vars\ (\be_1 \exor \be_2)\ \cid \defeq (\vars\ \be_1\ \cid) \cup (\vars\ \be_2\ \cid)
\end{mathpar}
... and so on. Then, to begin converting a protocol $\prot$ to a Datalog
we define the function ${tt}$ that applies $\solve{\!}{\!}{\!\!}$ locally to each command
in $\prog$, obtaining the subset of memories that result
in a variable assignment of $1$.  
\begin{definition} Define:
$$
\bodies(\xassign{x}{\be}{\cid}) \defeq (x, (\solve{(\mems(\vars\ \be\ \cid))}{\be}{\cid}))
$$
\end{definition}
The mapping of $\bodies$ across a program
$\prog$-- i.e., $(\mathit{map}\ \bodies\ \prog)$--  essentially defines the
logic program, modulo some syntactic conversion. We can
accomplish the latter as follows, where $\datalog(\prog)$ defines the
full conversion.
\begin{definition} We define the conversion from memories to
  literals and clause bodies as follows:
\begin{mathpar}
  \logit{x \mapsto 1} \defeq x

@@

  \logit{\{ x_1 \mapsto \beta_1, \ldots, x_n \mapsto \beta_n\}} \defeq
  \logit{x_1 \mapsto \beta_1} \wedge \cdots \wedge \logit{x_n \mapsto \beta_n}
\end{mathpar}
Given pairs $(x,\stores)$ in the range of $\bodies$, we define the conversion
to clauses as  $\mathit{clauses}(x,\{ \store_1,...,\store_n \}) \defeq x \gets \logit{\store_1} \vee \cdots \vee x \gets \logit{\store_n}$.
The $\minifed$-to-Datalog conversion is then defined as:
$$
\datalog(\prog) \defeq  \mathit{map}\ \mathit{clauses}\ (\mathit{map}\ \bodies\ \prog)
$$
\end{definition}

@@

\paragraph{Verifying Security Properties} We can
query correct representations of $\runs(\prog)$ by using
implementations of conditioning and marginalization to automatically
verify the passive model hyperproperties described in Section
\ref{section-hyper}-- in particular correctness and noninterference
modulo output, which imply passive security.  We have used this method
to verify security in $\mathbb{F}_2$ of Shamir addition as defined in Section
\ref{section-lang-example}, and single-gate $\eand$ and $\exor$ circuits with GMW
Beaver triples. We have also
verified properties compositional properties of gates themselves that establish
circuit invariants that imply passive security in larger circuits.
We discuss these examples and proof methods in Section
\ref{section-example-gmw}.

### 2024-05-14 ppdp24/example-bdoz.tex
checkpoint

\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$,
  let $X_H \subseteq \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$.
  Then the \emph{adversarial inputs to $X_H$} is the least set
  $X_C \subseteq \cinputs$ such that $\progtt(\prog) \not\vdash X_C * X_H$.
\end{definition}
\begin{definition}[Cheating Detection]
  \emph{Cheating is detected} in $\prog$ with $\iov(\prog) = S \cup V \cup O$ iff
  for all  $\store \in \aruns(\prog)$,
  letting $X_H = \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$,
  and letting $X_C$ be the adversarial inputs to $X_H$,
  there exists $\sigma'\in \runs(\prog)$
  with $\store_{X_C} = \store'_{X_C}$.  
\end{definition}

\begin{lemma}
  If cheating is detected in $\prog$, then $\prog$ has integrity.
\end{lemma}

>>>>

We can carry out similar proofs of passive security for the protocol in
Figure \ref{fig-beaver} as for GMW, even using automated tactics for
the protocol in $\mathbb{F}_2$. But in the case of BDOZ we are also
concerned with malicious security. To demonstrate this, we can
demonstrate that the protocol satisfies integrity in the sense of
Definition \ref{def-integrity}. To do so, we observe that it satisfies
a stronger property, that we call cheating detection. Intuitively,
integrity says that the only thing that the adversary can do in the
malicious model is to elicit the same responses from honest parties
that an honest run of the protocol would elicit. Cheating detection
says that the adversary can only execute the protocol honestly, or
or else gets caught (and abort).

Focusing in, we identify the adversarial inputs as the messages
sent from the adversary to honest parties, on which honest responses
to the adversary may depend. We want to say that these are the messages
that must be legitimate.
\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$,
  let $X_H \subseteq \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$.
  Then the \emph{adversarial inputs to $X_H$} is the least set
  $X_C \subseteq \cinputs$ such that $\progtt(\prog) \not\vdash X_C * X_H$.
\end{definition}
Now, we can characterize protocols with cheating detection as those where
adversarial inputs to honest reponses must themselves be constructed honestly. 
\begin{definition}[Cheating Detection]
  \emph{Cheating is detected} in $\prog$ with $\iov(\prog) = S \cup V \cup O$ iff
  for all  $\store \in \aruns(\prog)$,
  letting $X_H = \{ x \mid x \in (\houtputs \cup O_H) \wedge x \in \dom(\store) \}$,
  and letting $X_C$ be the adversarial inputs to $X_H$,
  there exists $\sigma'\in \runs(\prog)$
  with $\store_{X_C} = \store'_{X_C}$.  
\end{definition}

It is straightforward to demonstrate that cheating detection has integrity,
since only the ``passive'' adversary can elicit a response from honest parties. 
\begin{lemma}
  If cheating is detected in $\prog$, then $\prog$ has integrity.
\end{lemma}

In the case of BDOZ, cheating detection is accomplished by the information-theoretic
security of the encryption scheme\cite{XXX}. Furthermore, the symmetry of
the protocol in Figure \ref{fig-beaver} ensures that both parties will authenticate
shares, so it is robust to corruption of either party.

### 2024-05-14 ppdp24/model.tex
checkpoint

We consider random tapes to be selected from a uniform distribution
modulo relevant pre-processing, e.g., Beaver triples for efficient
multiplication, and/or MACed share distributions as in SPDZ \cite{XXX}.
We will also assume that any possible initial secrets
are equally likely when formulating the program distribution.
We define $\runs(\prog)$ as the set of final memories resulting
from execution of $\prog$ given any initial memory, and
treat all elements of $\runs(\prog)$ as equally likely.
This establishes the basic program distribution that can be
marginalized and conditioned to quantify input/output
information dependencies. 
%In this
%setting, given a program $\prog$ with $\iov(\prog) = S \cup V$ and
%$\flips(\prog) = F$ we will consider all $\store \in \mems(S \cup V
%\cup F)$ such that $ \config{\store_{S \cup F}}{\prog} \redxs
%\config{\store_}{\varnothing} $ to be equally probable, establishing
%the basic distribution of the program. %From this, we can immediately
%derive the marginal distribution of $S \cup V$ to reason about
%dependencies between secrets and views.
\begin{definition}
  \label{def-progd}
  \label{definition-progd}
  Given $\prog$ with $\secrets(\prog) = S$ and $\flips(\prog) = R$ and
  pre-processing predicate $\preproc$ on memories, define:
  $$
  \begin{array}{c}
    \runs(\prog) \defeq \\
    \{ \store \mid \exists \store_1 \in \mems(R) . 
    \exists \store_2 . \preproc(\store_2) \wedge
    %(\dom(\store) = \iov(\prog) \cup R) \wedge
    (\store_1 \cap \store_2,\prog) \redxs (\store,\varnothing) \}
  \end{array}
  $$
  Then the \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
  $$
  \progtt(\store) =  1 / |\runs(\prog)| \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
  $$
  By default, $\preproc(\store) \iff \dom(\store) = S$, i.e.,
  the initial memory contains all input secrets in a uniform
  marginal distribution.

@@

It is necessary to add $\bot$ to the range of memories since
the possibility of abort needs to be reflected in adversrarial
runs of a protocol. We can define $\aruns(\prog,\adversary)$
as the ``prefix'' memories that result from possibly-aborting
protocols, but we also need to ``pad out'' the memories
of partial runs with $\bot$, as we defin in $\botruns(\prog,\adversary)$,
to properly reflect the contents of views and outputs even in case of abort. 
\begin{definition}
  \label{def-progd}
  \label{definition-progd}
  Given program $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$, and
  any assumed pre-processing predicate $\preproc$ on memories, define:
  $$
  \begin{array}{c}
    \aruns(\prog) \defeq \\
    \{ \store \mid \exists \store_1 \in \mems(R) . 
    \exists \store_2 . \preproc(\store_2) \wedge
    %(\dom(\store) = \iov(\prog) \cup R) \wedge
    (\store_1 \cap \store_2,\prog) \aredxs (\store,\varnothing) \}
  \end{array}
  $$
  and also define the following which pads out undefined views and outputs with $\bot$:
  $$
  \begin{array}{l}
    \botruns(\prog) \defeq \\
    \qquad \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid \\
    \qquad \phantom{\{} \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
  \end{array}
  $$
  Then the \emph{$\adversary$ distribution of $\prog$}, written $\progtt(\prog,\adversary)$, is
  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
  $$
  \progtt(\prog,\adversary) =  1 / |\botruns(\prog)| \ \text{if}\ \store \in \botruns(\prog), \text{otherwise}\ 0
  $$
\end{definition}

>>>>

We consider random tapes to be selected from a uniform distribution
modulo relevant pre-processing, e.g., Beaver triples for efficient
multiplication, and/or MACed share distributions as in SPDZ \cite{XXX}.
We will also assume that any possible initial secrets
are equally likely when formulating the program distribution.
We define $\runs(\prog)$ as the set of final memories resulting
from execution of $\prog$ given any initial memory, and
treat all elements of $\runs(\prog)$ as equally likely.
This establishes the basic program distribution that can be
marginalized and conditioned to quantify input/output
information dependencies. 
%In this
%setting, given a program $\prog$ with $\iov(\prog) = S \cup V$ and
%$\flips(\prog) = F$ we will consider all $\store \in \mems(S \cup V
%\cup F)$ such that $ \config{\store_{S \cup F}}{\prog} \redxs
%\config{\store_}{\varnothing} $ to be equally probable, establishing
%the basic distribution of the program. %From this, we can immediately
%derive the marginal distribution of $S \cup V$ to reason about
%dependencies between secrets and views.
\begin{definition}
  \label{def-progtt}
  \label{def-progd}
  \label{definition-progd}
  Given $\prog$ with $\secrets(\prog) = S$ and $\flips(\prog) = R$ and
  pre-processing predicate $\preproc$ on memories, define:
  $$
  \begin{array}{c}
    \runs(\prog) \defeq \\
    \{ \store \mid \exists \store_1 \in \mems(R) . 
    \exists \store_2 . \preproc(\store_2) \wedge
    %(\dom(\store) = \iov(\prog) \cup R) \wedge
    (\store_1 \cap \store_2,\prog) \redxs (\store,\varnothing) \}
  \end{array}
  $$
  By default, $\preproc(\store) \iff \dom(\store) = S$, i.e.,
  the initial memory contains all input secrets in a uniform
  marginal distribution. Then the \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
  $$
  \progtt(\store) =  1 / |\runs(\prog)| \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
  $$

@@

It is necessary to add $\bot$ to the range of memories since
the possibility of abort needs to be reflected in adversrarial
runs of a protocol. We can define $\aruns(\prog,\adversary)$
as the ``prefix'' memories that result from possibly-aborting
protocols, but we also need to ``pad out'' the memories
of partial runs with $\bot$, as we defin in $\botruns(\prog,\adversary)$,
to properly reflect the contents of views and outputs even in case of abort. 
\begin{definition}
  \label{def-progd}
  \label{definition-progd}
  Given program $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$, and
  any assumed pre-processing predicate $\preproc$ on memories, define:
  $$
  \begin{array}{c}
    \aruns(\prog) \defeq \\
    \{ \store \mid \exists \store_1 \in \mems(R) . 
    \exists \store_2 . \preproc(\store_2) \wedge
    %(\dom(\store) = \iov(\prog) \cup R) \wedge
    (\store_1 \cap \store_2,\prog) \aredxs (\store,\varnothing) \}
  \end{array}
  $$
  where by default, $\preproc(\store) \iff \dom(\store) = S$, and also define the following
  which pads out undefined views and outputs with $\bot$:
  $$
  \begin{array}{l}
    \botruns(\prog) \defeq \\
    \qquad \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid \\
    \qquad \phantom{\{} \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
  \end{array}
  $$
  Then the \emph{$\adversary$ distribution of $\prog$}, written $\progtt(\prog,\adversary)$, is
  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
  $$
  \progtt(\prog,\adversary) =  1 / |\botruns(\prog)| \ \text{if}\ \store \in \botruns(\prog), \text{otherwise}\ 0
  $$
\end{definition}
